---
title: "Bayesian Analysis Project"
author: "Maxime Jurado and Mathieu Marauri"
date: "Tuesday, May 26, 2015"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: no
    number_sections: no
    toc: yes
header-includes: \usepackage{float} \usepackage{graphicx}
---

```{r initial_chunk, echo = FALSE, warning = FALSE, message = FALSE}
library("knitr")
library("ggplot2")
library("xtable")
library("dplyr")
library("gridExtra")
library("xtable")
opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE)
options("xtable.comment" = FALSE)
options("xtable.table.placement" = "H")
```

```{r data}
data <- read.csv2("Data/data.csv",header=TRUE)
data <- subset(data,time!="TP")
data1 <- data[rep(1:nrow(data),each=5),]
data2 <- data.frame(data$X1,data$X2,data$X3,data$X4,data$X5)
data3 <- data.frame(data1$id,data1$time,data1$sexe)
vec <- NULL
for(i in 1:nrow(data2)){
  vec1 <- data2[i,]
  vec <- c(vec,t(vec1))
}
vec <- as.data.frame(vec)
data4 <- data.frame(data3,vec)
cat <- c(rep(c(1,2,3,4,5),132))
data <- data.frame(data4,cat)
colnames(data) <- c("id","time","sexe","salary","spc")
attach(data)
salary<-as.numeric(salary)
```

# Abstract

# Introduction

# Dataset
\label{dataset}

The dataset comes from the *Institut National de la Statistique et des Etudes Economiques* or INSEE in France. It s the statistical instuitute of France. The dataset contains 33 sectors of activity that are identified by the *id* variable. The classification can be found in table \ref{indexSector}. 

The variables reported in this study are listed below:

* Time is an indicator which takes value 1 for a full-time work and 0 for a part-time work. _time_
* Sexe is a binary variable with value 1 for male and 0 for female. _sexe_
* Socio-Professional Category: it takes values between 1 and 5. Table \ref{indexCat} shows the classification. _spc_
* Salary: it is the response variable. _salary_

The salary is the average gross income for an hour. For instance it can be the average income of female employees working full-time in the extractive industry.

TODO: something about the construction of the dataset (more than 5 per cell...)

# Descriptive analysis

An exploration of the data through a descriptive analysis gives first insights on the behavior of the variables. The point here is to see graphically if some varirables seem to have an impact on the response variable, the salary.

## The Socio-Professional Categories

The first idea is that the salary is supposed to be different based on the different Socio-Professional Categories. The following graph (Figure \ref{BoxplotCat}) shows the mean of the salary by SPC.
```{r boxplotcat, fig.cap="Boxplot of the salary for each category. \\label{BoxplotCat}", fig.pos="H", fig.align='left'}
theme <- theme(plot.title = element_text(size=10, face="bold"), axis.title.x = element_text(size=10), axis.title.y = element_text(size=10), axis.text.x = element_text(size=10, face="bold"),axis.text.y = element_text(size=10, face="bold"))
ggplot(data, aes(factor(spc), salary)) + geom_boxplot(aes(fill = factor(spc))) + xlab("Socio-Professional Category") + ylab("Salary") + guides(fill=guide_legend(title=NULL)) + theme
```

One can clearly see that being in the category 1 (Higher managerial and professional positions) increases the salary as compared to the other categories. If one considers the categories to be ordered from 1 to 5 then he would be able to say that _SPC_ seems to have a negative impact on the salary.

In order to see more clearly the way _SPC_ influences the salary Figure \ref{completeplot} shows the evolution of the salary for each SPC for each sector. The means were calculated on all the values of salary for the SPC by id.
```{r completeplot, fig.cap="Evolution of the salary for each SPC by id. \\label{completeplot}", fig.pos="H", fig.align='left'}
datamean <- aggregate(data.frame(salaryMean = data$salary), by = list(id = data$id, spc = data$spc),mean)
ggplot(datamean, aes(spc, salaryMean)) + facet_wrap(facets=~ id, ncol=10) + geom_line(colour="blue") + geom_point(colour="red") + theme
```

As seen previously _SPC_ seems to have a negative impact on the salary for every sectors but sector 6 (Manufacture of food, drink and tabacoo based products), sector 21 (Information and communication) and sector 30 (Arts and entertainment). This may be due to the definitions of the SPC. In most of the sectors they are ordered from 1 to 5 but in some other sectors this order can be different. 

## The time

TODO: salary per hour definition


```{r boxplottime, fig.cap="Boxplot of the salary for each time. \\label{boxplottime}", fig.pos="H", fig.align='left'}
ggplot(data, aes(factor(time), salary)) + geom_boxplot(aes(fill = factor(time))) + labs(x="Time", y="Salary", title="Boxplot of the salary for each time") + theme + scale_fill_manual(name="Time", values=c("orange", "mediumpurple"), labels=c("0"="Part-time", "1"="Full-time"))
```



```{r boxplotspctime, fig.cap="Boxplot of the salary for each SPC and each time. \\label{boxplotspctime}", fig.pos="H", fig.align='left'}
ggplot(data, aes(factor(spc), salary)) + geom_boxplot(aes(fill = factor(time))) + labs(x="SPC", y="Salary") + theme + scale_fill_manual(name="Time", values=c("orange", "mediumpurple"), labels=c("0"="Part-time", "1"="Full-time"))
```


# Statistical methods

# Bayesian model 

# Bayesian model with random effects

# COnclusion

# Bibliographie

# Appendix

## Classification tables

Here are presented the classification tables for the Socio-Professional Categories (Table \ref{tableIndexSpc}) and for the Sectors (Table \ref{tableIndexSectors}).

```{r classification table spc, results='asis'}
categories <- read.table("Data/Index SCP.txt", sep="-", header=FALSE)
categories <- data.frame(categories$V2)
colnames(categories) <- "Category"
print(xtable(categories, align=c("c","c"), caption="Classification table of the Socio-Professional Categories. \\label{tableIndexSpc}"))
```

```{r classification table sectors, results='asis'}
sectors <- read.table("Data/Index Sectors.txt", sep=";", header=FALSE)
sectors <- data.frame(sectors$V2)
colnames(sectors) <- "Sector"
print(xtable(sectors, align=c("c","c"), caption="Classification table of the Sectors. \\label{tableIndexSectors}"))
```

Reurn to section \nameref{dataset}.
